---
title: "Untitled"
author: "Mustafa Telab"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(plotly)
library(shiny)
library(tidyverse)
```


Module 3
I have provided you with data about mortality from all 50 states and the District of Columbia.
Please access it at
https://github.com/charleyferrari/CUNY_DATA608/tree/master/module3/data You are
invited to gather more data from our provider, the CDC WONDER system, at
https://wonder.cdc.gov/ucd-icd10.html .
● This assignment must be done in R. It must be done using the ‘shiny’ package.
● It is recommended you use an R package that supports interactive graphing such as
plotly, or vegalite, but this is not required.
● Your apps must be deployed, I won’t be accepting raw files. Luckily, you can pretty
easily deploy apps with a free account at shinyapps.io

```{r,message=FALSE}
#IMPORT DATA
df = read_csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")

```

```{r}
#CALCULATE RATE
df['rate'] = round((df['Deaths']/df['Population'])*1000,digits = 3)
#CREATE UNIQUE LISTS FOR DROP DOWN OPTIONS
causes = unique(df[1])
geos = unique(df[2])
```

Question 1:
As a researcher, you frequently compare mortality rates from particular causes across
different States. You need a visualization that will let you see (for 2010 only) the crude
mortality rate, across all States, from one cause (for example, Neoplasms, which are
effectively cancers). Create a visualization that allows you to rank States by crude mortality
for each cause of death.

```{r}
ui <- fluidPage(
selectInput(inputId = "causes",label = "Select Mortality",choices = causes ),
plotOutput('hist')
)
server = function(input,output){
  output$hist = renderPlot({
dfsub%>%
      filter(ICD.Chapter == input$causes)%>%
          ggplot(aes(x = reorder(State,Crude.Rate), y = Crude.Rate)) + 
            geom_bar(stat = 'identity') +
            xlab("")+
            ylab("Deaths")+
            coord_flip() 
  })
}
shinyApp(ui = ui, server = server)
```

Question 2:
Often you are asked whether particular States are improving their mortality rates (per cause)
faster than, or slower than, the national average. Create a visualization that lets your clients
see this for themselves for one cause of death at the time. Keep in mind that the national
average should be weighted by the national population.

```{r}
ui <- fluidPage(
selectInput(inputId = "causes",label = "Select Mortality",choices = causes ),
selectInput(inputId = "State",label = "Select State",choices = geos),
plotOutput('hist')
)
server = function(input,output){
  output$hist = renderPlot({
    df%>%
      filter(ICD.Chapter == input$causes, State == input$State)%>%
          ggplot(aes(x = Year, y = Crude.Rate)) + 
            geom_line(stat = 'identity') +
            xlab("")+
            ylab("Deaths")
  })
}
shinyApp(ui = ui, server = server)
```

